---
title: "CourseProject1"
author: "David Holtzhouser"
date: "February 5, 2016"
output: html_document:
  keep_md: yes
---

### Install and load the required packages:
```{r tidy=TRUE}
  r_repos <- "http://cran.rstudio.com/"
  install.packages("RCurl", repos = r_repos)
  install.packages("rmarkdown", repos = r_repos)
  install.packages("data.table", repos = r_repos)
  install.packages("ggplot2", repos = r_repos)
  install.packages("lattice", repos = r_repos)

  library(RCurl)
  library(markdown)
  library(data.table)
```

***

###  1) Get and clean the activity data:
```{r tidy=TRUE}
  temp <- tempfile()
  gitHubUrl <- "https://github.com/dholtz/RepData_PeerAssessment1/blob/master/activity.zip?raw=true"
  download.file(gitHubUrl, temp)
  activityMonitoringData <- read.csv(unz(temp,"activity.csv"))
  unlink(temp)
  cleanActivityMonitoringData <- na.omit(activityMonitoringData)
```


***

###  2) Total Number of Steps Taken Each Day

#### 2a) Total Number of Steps taken per Day
```{r tidy=TRUE}
  stepsPerDay <- aggregate(steps ~ date, cleanActivityMonitoringData, sum)
```

#### 2b) Total Number of Steps taken each Day (Historgram)
```{r tidy=TRUE}
  hist(stepsPerDay$steps, main = "Total Steps Taken per Day", xlab = "Steps Per Day", col = 1)
```

#### 2c) Mean and Median of total number of steps taken per day
```{r tidy=TRUE}
  mean(stepsPerDay$steps)
  median(stepsPerDay$steps)
```


***

### 3) What is the average daily activity pattern?
```{r tidy=TRUE}
  intervalOfStepsAverage <- aggregate(steps ~ interval, cleanActivityMonitoringData, mean)
```

#### 3a) Time Series plot
```{r tidy=TRUE}
  plot(intervalOfStepsAverage$interval, intervalOfStepsAverage$steps, type='l', main = "Average Daily Activity Pattern",
     xlab = "5 Minute Time Intervals", ylab = "Average Number of Steps")
```

### 3b) 5-minute interval, on average across all days, contains the maximum number of steps?
```{r tidy=TRUE}
  maxIntervalSteps <- intervalOfStepsAverage[which.max(intervalOfStepsAverage$steps), ]
```
#### -The interval, `r maxIntervalSteps$interval` contained, `r maxIntervalSteps$steps` steps.


***

###  4) Strategy for imputing Missing Data

#### 4a) Total number of missing values in the dataset (the total number of rows with NAs)
```{r tidy=TRUE}
  totalNumNAs <- colSums(is.na(activityMonitoringData))
```
##### - There are `r totalNumNAs["steps"]`, NA's in the steps, `r totalNumNAs["date"]`, NA's in the date,
##### and `r totalNumNAs["interval"]`, NA's in the interval.

***

###  5) Total Number of Steps Taken Each Day, AFTER missing values are imputed

#### 5a) Fill in the missing values with the mean value for the 5-minute interval
```{r tidy=TRUE}
  mean5MinIntervalFn <- function(steps, interval) {
    filled <- NA
    if (!is.na(steps)) {
      filled <- c(steps)
    } else {
      filled <- intervalOfStepsAverage[intervalOfStepsAverage$interval == interval, "steps"]
    }
    filled
  }
  
  missingValuesFilled <- activityMonitoringData
  missingValuesFilled$steps <- mapply(mean5MinIntervalFn, missingValuesFilled$steps, missingValuesFilled$interval)
  missingFilledIntervalOfStepsAverage <- aggregate(steps ~ date, missingValuesFilled, sum)
```

```{r tidy=TRUE}
  #Sanity check that all NAs are removed from original data set
  postNAFillCheck <- colSums(is.na(missingValuesFilled))
```
##### - There are `r postNAFillCheck`, NA's in the filled data set.

#### 5b) Compare the original steps taken each day with NA filled in data.
```{r tidy=TRUE}
  par(mfrow=c(2,1))
  
  hist(stepsPerDay$steps, main = "Total Steps Taken per Day / Original Data", xlab = "Steps Per Day", col = 1)
  hist(missingFilledIntervalOfStepsAverage$steps, main = "Imputed Total Steps Taken per Day / NA's Removed Data", xlab = "Steps Per Day", col = 1)
  
  mean(stepsPerDay$steps)
  median(missingFilledIntervalOfStepsAverage$steps)
```  

#### - There are NO differences from the initial estimates from the first part of the assignment after imputing missing data.

***

###  6) Comparing the Average Number of Steps taken per 5-minute Interval Across Weekdays and Weekends

#### 6a) Create factor variable in the 'filled' data set with two levels, "weekday" and "weekend" indicating
#### whether a given date is a weekday or weekend day.
```{r tidy=TRUE}
  weekDayOrWeekEnd <- function(strDate) {
    w <- NA
    val <- weekdays(as.Date(strDate, '%Y-%m-%d'))
    if (!(val == 'Saturday' || val == 'Sunday')) {
      w <- 'Weekday'
    } else {
      w <- 'Weekend'
    }
    w
  }
  
  missingValuesFilled$typeOfDay <- as.factor(sapply(missingValuesFilled$date, weekDayOrWeekEnd))
  
  imputedStepsPerDay <- aggregate(steps ~ interval+typeOfDay, missingValuesFilled, mean)
  
  #load ggplot 2 library
  library(ggplot2)

  qplot(interval, steps, data = imputedStepsPerDay, geom = c("line"), xlab = "5 Minute Time Intervals",
       ylab = "Average Number of Steps", main = "Number of steps Per Interval by Day Type") + 
      facet_wrap(~ typeOfDay, ncol = 1)

```

##### - From the plot, it appears that the person is more active on the weekends , with a slight peak during the week during the morning hours.

***

#### - This can be more easily visualized by showing the data on a single plot.
```{r tidy=TRUE}
  library(lattice)
  
  xyplot(steps ~ interval, data = imputedStepsPerDay, groups = imputedStepsPerDay$typeOfDay, type="l", grid=T,
         ylab="Average number of steps", xlab="5-min. intervals from midnight", 
         main="Weekdays (in blue) vs. Weekends (in Purple)") 
```
