---
title: "CourseProject1"
author: "David Holtzhouser"
date: "February 5, 2016"
output: html_document
---

### Install and load the required packages:
```{r}
r_repos <- "http://cran.rstudio.com/"
install.packages("RCurl", repos = r_repos)
install.packages("rmarkdown", repos = r_repos)
install.packages("data.table", repos = r_repos)

library(RCurl)
library(markdown)
library(data.table)
```


###  1) Get and clean the activity data:
```{r}
temp <- tempfile()
gitHubUrl <- "https://github.com/dholtz/RepData_PeerAssessment1/blob/master/activity.zip?raw=true"
download.file(gitHubUrl, temp)
activityMonitoringData <- read.csv(unz(temp,"activity.csv"))
unlink(temp)
cleanActivityMonitoringData <- na.omit(activityMonitoringData)
```

###  2) Total Number of Steps Taken Each Day

#### 2a) Total Number of Steps taken per Day
```{r}
stepsPerDay <- aggregate(steps ~ date, cleanActivityMonitoringData, sum)
```

#### 2b) Total Number of Steps taken each Day (Historgram)
```{r}
hist(stepsPerDay$steps, main = "Total Steps Taken per Day", xlab = "Steps Per Day", col = 1)
```

#### 2c) Mean and Median of total number of steps taken per day
```{r}
mean(stepsPerDay$steps)
median(stepsPerDay$steps)
```

### 3) What is the average daily activity pattern?
```{r}
intervalOfStepsAverage <- aggregate(steps ~ interval, cleanActivityMonitoringData, mean)
```

#### 3a) Time Series plot
```{r}
plot(intervalOfStepsAverage$interval, intervalOfStepsAverage$steps, type='l', main = "Average Daily Activity Pattern",
     xlab = "5 Minute Time Intervals", ylab = "Average Number of Steps")
```

### 3b) 5-minute interval, on average across all days, contains the maximum number of steps?
```{r}
  intervalOfStepsAverage[which.max(intervalOfStepsAverage$steps), ]
```



###  6) Strategy for imputing Missing Data



###  7) Total Number of Steps Taken Each Day, AFTER missing values are imputed (Histogram)


###  8) Comparing the Average Number of Steps taken per 5-minute Interval Across Weekdays and Weekends




###  9) R Code to reproduce these results in the report (numbers, plots, etc.)



You can also embed plots, for example:

```{r, echo=FALSE}
plot(cars)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
